---
title: "Samp"
subtitle: "Applied Machine Learning"
author: "Jameson > Hendrik > Calvin"
        
revealjs-plugins:
  - drop

---


## Agenda

1. Course Overview
2. Review of Regression
3. Dinner Break
4. Classification and Ethics
5. Basic Feature Engineering
6. Vocabulary

# Course Overview

## Expectations and assignments

1. Homework assignments
2. Exams
3. Modeling Project
4. Course Policies
5. My expectations for you

## About me
- BA Mathematics, BS Computer Science (UChicago)
- MS, PhD Computer Science (UNC Chapel Hill)
- Data mining, formal analysis, complex models
- Joined Willamette 2021

## About you?
  - Background
  - Goals for this program and/or course

## Basic concepts in Machine Learning

- What is a data scientist?
- What is machine learning? 
- What is the role of judgment in machine learning?
- What are the differences between machine learning, statistics and econometrics?
- When is "mere" correlation enough? When is it not?

## Packages

- Today I use the following libraries:
```r
local({r <- getOption("repos")
       r["CRAN"] <- "https://cran.r-project.org" 
       options(repos=r)
})
# New?
install.packages("tidyverse")
install.packages("moderndive")
install.packages("caret")
install.packages("dslabs")
# Just for the slides
install.packages("thematic")
```
- You will have some but perhaps not others.

## Libraries

- I'll just include them upfront.
```{r Libraries}
library(tidyverse)
library(moderndive)
library(caret)
library(dslabs)
# Just for the slides
library(thematic)
theme_set(theme_dark())
thematic_rmd(bg = "#111", fg = "#eee", accent="#eee")
```

## Partition our Data

```{r Partition}
data(heights) # from library(dslabs)
y <- heights$sex
x <- heights$height
set.seed(505)
test_index <- createDataPartition(y, times = 1, p = 0.5, list = FALSE)
test_set <- heights[test_index, ]
train_set <- heights[-test_index, ]
summary(heights)
```
